Leadership across cultures and contexts: an exploratory analysis

### 2020-04-07 Response to reviewers.Rmd

We now discuss these limitations and points more clearly in the limitations section. NEED TO DO.

**What does it mean for someone to be a leader at the "supracommunity" level? I'm not familiar with the term, so had difficulty contextualizing the relevant findings. And how exactly do the authors define "community" versus "supracommunity"?**

In the methods section we now include a description of all levels of our *group structure* coding to clearly explain this coding scheme. We coded supracommunity leaders in cases when the "group" involves leadership beyond the community level and across multiple residential communities, such as regional political leaders, chiefs in complex chiefdoms, kings, and state-level leaders. This termonology was vaugely inspiried by Murdock and Provost's (1971) Political Integration Scale which codes the political integration at the society-level as either automonous local communities, or having one, two, or three levels of juridictional hiuerarchy "above the community"; i.e., if political authority is exclusive to a local village headman, the group would be coded as "autonomous" but if the village and headman are under authority of a chief from a seperate community, the group would be coded as "one level above community", and so forth.

### 2020-04-07 Response to reviewers.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Response to editor and reviewer comments

### Editor comments

**Statistical reporting and further analysis: The GLMM should be reported in full somewhere. This could go in the supplemental materials. However, can we learn more from it? Do the random effects matter for the ethnographers? What about the national origin effects for the ethnographers? We recommend taking seriously the reflexive analytical techniques of putting the ethnographer in the analysis (see work by Robert Aunger working in the Congo Basin). Does the ethnographer matter? If so, how much?**

**Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population? This can (potentially) speak quantitatively to some of the qualitative discussion about limitations and biases in the approach.**

We determined the gender of almost all authors of all documents included in our study. We found that of the 884 text records with no female author or co-author, 21 (2.4%) reported female leadership. Of the 225 text records with a female author or co-author, 20 (8.9%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = 3.5), it was not statistically significant (p = 0.25). We now report this in our results section.

@@

**Finally, the reviewers made a number of valuable suggestions, not enumerated above,  that should help you sharpen you presentation and discussion. Please attend to those.**

## Reviewer 1 comments

**This manuscript presents a descriptive analysis of the HRAF to provide information about leadership in traditional societies. Overall, I enjoyed reading the paper and thought it was a useful resource. It provides a comprehensive analysis of leadership across a wide swath of traditional cultures. It reports on the types of functions leaders serve, the qualities possessed by leaders, and the costs and benefits of leadership experienced by both leaders and followers. The paper will likely be of interest to anyone working in the area of leadership, social hierarchy, or group behavior. I provide some comments below aimed at helping the authors increase the potential accessibility and impact of the work.**

We greatly appreciate the reviewers detailed comments and very useful suggestions.

**The paper was framed partially as providing an analysis of differences between traditional societies and WEIRD societies but, unless I missed it, there was not much analysis along these lines. Neither the background section nor the Discussion section focuses much on contrasting WEIRD and traditional societies. The paper appears to miss out on a useful opportunity to provide some commentary on this issue. Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**

We agree our initial submisson did not sufficently accomplish this goal. We have now added summarizing, concluding statements on the qualities and functions of leaders in WEIRD and non-WEIRD settings in the introduction.

@@

**The abstract could be tightened up in terms of description of particular costs and benefits to leaders and followers. And a brief justification in the abstract for why the study compares costs and benefits would be useful- e.g. understanding evolution of leader or follower strategies requires understanding the (fitness-relevant) costs and benefits to them.**

**Lines 88-89 suggest that the paper's ethnographic analysis will include leadership at all levels of social organization. But does the theory and analysis reported in the paper really extend to all levels of social organization- such as child play groups or households? It seems the focus is still community-level cooperation among adults. Perhaps clarify.**

**In the "Functions of leadership" section, I would expand on your claim that leadership emerges when there is demand for a "managerial presence". You might describe collective action or coordination problems here. That might better motivate the subsequent discussion of different functions of leadership in small-scale societies, as well as discussion of the conditions (e.g. demographic change, wealth access) that increase institutionalization or authority of leaders. Glowacki and von Rueden (2015) suggest that inter-family conflict resolution is where leader institutionalization often first arises, and Roscoe (2009) makes a similar argument concerning leadership in highland New Guinea, where the institutionalized display of material wealth organized by Big Men is rewarded more than ad hoc leadership during war.**

**The Wofford study (lines 214-233) and GLOBE study reminds me of the vast array of descriptive models in leadership/managerial studies- transactional vs. transformational leadership and their sub-components being one of the most popular. So what should a reader make of this literature, with respect to evolution of leadership in small-scale societies? Do bureaucracies, formal legal systems, formal education-based skills, and large inequalities in remuneration limit universality of descriptive models from leadership/managerial studies? I'd be more explicit on this in the introduction. For the discussion section: how does your own descriptive modeling of leadership in traditional societies agree or not with Wofford or other models from large-scale societies?**

### 2020-04-09 Response to reviewers.Rmd

## Response to editor and reviewer comments

### Editor comments

**Statistical reporting and further analysis: The GLMM should be reported in full somewhere. This could go in the supplemental materials. However, can we learn more from it? Do the random effects matter for the ethnographers? What about the national origin effects for the ethnographers? We recommend taking seriously the reflexive analytical techniques of putting the ethnographer in the analysis (see work by Robert Aunger working in the Congo Basin). Does the ethnographer matter? If so, how much?**

Although we determined the gender of almost all ethnographers (see below), we were not able to confirm the national origins of most.

We now briefly describe the random effects for ethnographers. [need to do]

@@

We determined the gender of almost all authors of all documents included in our study. We found that of the 884 text records with no female author or co-author, 21 (2.4%) reported female leadership. Of the 225 text records with a female author or co-author, 20 (8.9%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = 3.5), it was not statistically significant (p = 0.25). We now report this in our results section.

**Could further analyses be done to obtain more information on the nature of the variation within the data?**

**1.	The cluster analysis is nice, but it opens up the option to combine items within both qualities and functions to create a smaller number of higher-level categories or codes that, arguably, are better measured. The issue here is that some things that don’t appear frequently might rise in frequency when they are combined with other similar items. Suppose you use a higher-level category that contains 4 codes and you code the society as ‘1’ if any of the codes appears and zero otherwise. In our view, you need not go crazy with this approach, but look for any tightly bound clusters and consider pursuing those. If you are able to do this, the presentation could be more compact and easier to digest for the reader.**

In the original version of the ms we did not describe our numerous efforts at dimension reduction, including standard PCA, various flavors of logistic PCA, non-negative matrix factorization (NMF), and various flavors of cluster analysis, and we regressed the resulting components against our various culture-level variables. These efforts did not yield any convincing "stucture" in our data or results beyond what we reported. We now briefly describe these efforts so as to better contexualize the results that we do report.

We now additionally create variables based on the clusters we did find, and investigate their relationship to culture-level variables. As before, we find little convincing evidence of meaningful patterns, which we now briefly report in the SI.

@@

**2.	Why not take either some of your original codes or some of these higher-level codes (from the above cluster analysis) and regress them on societal-level features related to subsistence type (foragers, pastoralist, etc. or simply forager vs. non-forager if that’s appropriate), levels of jurisdictional hierarchy (or other measures of societal complexity), population size, and potentially ecological variables. Your sample is not big, but you might be able to get some juice out of this. If interesting findings do emerge, you should consider adding or otherwise altering your visualizations.**

See above.

**3.	Be clearer about how you are accounting for research effort in each society. For example, are you controlling for the total number of words or pages for each population?**

Total number of pages of ethnography for each culture. Results now reported.

**This seems very important since we want to make sure that the variation observed isn’t due to some populations getting more or less total study effort or writing effort.**

We investigated three possible sources of bias: total number of pages of ethnography, year of publication of each document, and the presence or absence of a female author/co-author for each document. We found evidence for bias in reporting of evidence for 9 of 105 [get new value] variables based on year of publication only. There was no convincing evidene of bias in any variable based on the total number of pages of ethnography or ethographer gender. Results now reported.

@@

**4.	Did you control for observation year? This might be too highly correlated with ethnographer identity, but worth checking**

Yes. See above.

**5.	Phylogenetic controls. While this is the probability sample, worries persist about non-independence (see R1). There are a variety of approaches to mitigate (and study!) this. Is there any clumping among related groups that cannot be explained by other variables? Do the patterns hold when you either control for linguistic distance or cluster in some way (continental fixed effects at a minimum)?**

We agree that worries persist about the putative independence of the PSF. We now investigate continental effects for each variable. We find xxx, yyy, zzz.

@@

**6.	Analyze the degree to which ‘researcher efforts’ (say measured by total words on that society and/or number of paragraphs on leadership) predict the appearance of a code. For example, what if the total ethnographic pages strongly predict the the appearance of “high status” as a code. Suppose the 10% of populations who don’t mention “high status” hardly wrote anything about leadership. On the other hand, if for other traits, there’s no relationship between researcher effort and appearance, then a case can be made that absence of evidence could very well be evidence of absence.**

As discussed above, we investigated the possible effects of total number of pages of ethnography for each culture on each variable. As we now report, we found no significant effects.

@@

**We recommend making liberal use of supplemental materials. Some of the analyses may not yield findings of sufficient interest to warrant placement in the main text, but some readers may want to have a look at them, so put them in the supplemental and briefly mention them in the main.**

We have expanded our supplemental materials, xx, yy, zz.

**Returning to the WEIRD Problem: in your discussion, please circle back to the question of what research might be missing by focusing on leadership in Western or industrialized societies. Say more on how leaders in small-scale societies are the same and different—and perhaps you will be able to explain some of that variation with variables like jurisdictional hierarchy or subsistence type or mobility measures. As mentioned by R1, this likely involves returning to the GLOBE data and putting it together with your findings more directly.**

We agree we need to better situate our study within the theme of the special issue. We interpret "Beyond WEIRD" in the following way: The original WEIRD paper was a needed wake-up call for psychology. However, we now worry that many psychologists are operationalizing WEIRD as a dichotomy: WEIRD and non-WEIRD. This is obviously a mistake, as there is enormous variation among the non-WEIRD. Our data set focuses primarily on traditional societies, and thus speaks to variation among the non-WEIRD, a point we now emphasize. At the same time, as Edward Said famously argued, anthropologists tend to make the opposite mistake: exoticizing the "Other". We hope we have struck the right balance in identifying important common themes and patterns across most of the populations in our data, while doing justice to their differences.

@@

**Status vs. Leadership: R2 asks about status vs. leadership. Given recent debates about a “leadership psychology” as distinct and separable from status psychologies (like prestige and dominance, van Vugt), it might be worth expanding a bit on what these data say about the connection between status and leadership.**

This is a good idea. We now include a short section on our "status" variable, and what distinguishes leadership records with evidence for social status from those without such evidence.

@@

**Expanding the dataset: the natural next step with this project is to use this small sample of hand coded categories to create an automated approach for all of the eHRAF, perhaps based on key words. The idea is to develop a set of keywords or other patterns that reliably predict when a paragraph should be coded in a particular way. The method can be developed and benchmarked using the dataset you’ve created and then simply applied in automated fashion to the entire eHRAF corpus. We offer this as a suggestion, but realize that it’s a lot to ask in this context.**

This is a good suggestion for future research, but goes beyond remit of the current paper.

### 2020-04-09 leadership_across_cultures_contexts.Rmd
Changes to all docs. Using margins package. More responses to reviewers. Removed vars with v. little evidence. Supplementary info.

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture. The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(leader_text2)` text records, `r final_record_count` had evidence for at least one variable in the study.

@@

@@

```{r cluster, fig.cap="Cluster analysis of leader quality variables (left) and leader function variables (right). Distances are $1 - cor$. Ward agglomeration method. AU p-values (red) computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
# plot(m_pvclust_qual)
# pvrect(m_pvclust_qual, alpha = 0.9)
knitr::include_graphics("Figures/m_pvclust.both.png")
```

@@

## Effects of document meta-data on leadership patterns

We explored whether document features, such as date of publication and gender of authors, influenced reporting of leader functions and qualities. There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine, some of which were co-authors. We then tested if text records with a female author or co-author were more likely to discuss female leaders. Of the `r sum(tab_coauthor['FALSE',])` text records with no female author or co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership. Of the `r sum(tab_coauthor['TRUE',])` text records with a female author or co-author, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).

@@

@@

>>>>

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture. Our search yielded `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures. The number of text records per culture ranged from `r min(dfwc$number_leader_records, na.rm=T)` to `r max(dfwc$number_leader_records, na.rm=T)`, with a median of `r median(dfwc$number_leader_records, na.rm=T)`, a mean of `r signif(mean(dfwc$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(dfwc$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(leader_text2)` text records, `r final_record_count` had evidence for at least one variable in the study.

@@

## Possible sources of bias

We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)). Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)`. There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors). Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author. Total pages of ethnography on each culture in the PSF ranged from `r min(dfwc[['N Pages eHRAF']])` to `r as.integer(max(dfwc[['N Pages eHRAF']]))`, with a median of `r median(dfwc[['N Pages eHRAF']])`.

We fit `r length(all_study_vars)` logistic mixed effects models with each binary study variable as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures. We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995]. After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models. More recent publications were more likely to report evidence for Oratory skill, Group representative, Group determination/cohesiveness, and Knowledgeable/intelligent, and less likely to report evidence for leaders imposing costs (Punishment, Killer) or receiving benefits (Social services, Food, and Reduced risk of harm). See Figure \@ref(fig:pubdate). See supplementary information for details.

We then tested if text records with a female author or co-author were more likely to discuss female leaders. Of the `r sum(tab_coauthor['FALSE',])` text records with no female author or co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership. Of the `r sum(tab_coauthor['TRUE',])` text records with a female author or co-author, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).

@@

```{r cluster, fig.cap="Cluster analysis of leader quality variables (A) and leader function variables (B). Distances were $1 - cor$. Ward agglomeration method. AU p-values (red) computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
# plot(m_pvclust_qual)
# pvrect(m_pvclust_qual, alpha = 0.9)
knitr::include_graphics("Figures/m_pvclust.both.png")
```

@@

@@

## Possible ethnographic or publication biases

@@

We explored three possible sources of bias in discovery of evidence for each of our `r length(all_study_vars)` variables: publication date, female vs. male-only authorship, and variation in the total number of pages of ethnography for each culture. For each binary study variable, we fit a logistic mixed effects model with that variable as the outcome, the bias variables as fixed effects, and random intercepts for ethnographers nested within cultures. We fit two sets of models. The first univariate set fit separate models for each study variable and each bias variable (total number of models = 105x3 = 315). The second multivariate set included all bias variables in each model (total number of models = 105). All p-values were corrected for a FDR of 5%. After correction, the coefficients of female authorship and pages of ethnography were not significant in any model. Seven variables had significant associations with year of publication in both the univariate or multivariate sets (with nearly identical coefficients), but Group determination/cohesiveness was only significant in the univariate model, and Leader social service benefits was only significant in the multivariate model. See Figure \@ref(fig:pubdate).

```{r pubdate, fig.cap="The effect of publication date on the probability of finding evidence for leader qualities, functions, benefits, and costs. Multivariate values are  controlling for female authorship and number of pages of ethonography for each culture. All models had random intercepts for ethnographers nested within cultures. Models were fit for each of the 105 variables as outcomes, and after adjusing p-values for a false discovery rate of 5%, the effect of publication date on these outcome variables remained significant. Publication date was centered and standardized by 1 SD. Each dot is one coefficient (converted from log odds to odds ratios). Bars are 95% CI."}
plot_pubdate
```

## Feature models

```{r features, fig.width=10, fig.cap="Marginal effects of Poission mixed effects regression models of Feature variables by mode of subsistence. Only models whose subsistence coefficient had a corrected p-value < 0.05. Base level: agriculturalists. See text for details."}

x <- pmap(sub_models_sig, ~ plot(..6, main = ..1))

```
